---
title: "Applying Augur to PBMC data"
author: "P. Nicol"
date: "8/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Summary** In this file I apply Augur to the PBMC dataset. You'll need to have Augur installed (do this from GitHub) to run this code. 


## Load the data 

As always, we load the data 

```{r}
Sco <- readRDS("/Users/phillipnicol/Desktop/local_files/data/scDiffPopDATA/gse145281/seurat.RDS")
Sco <- subset(Sco, idents = c(15, 16), invert = TRUE)
```

We need to adjust the metadata so that Augur will accept it.

```{r}
cell_types <- c("Naive CD4+ T", "CD36+ Monocytes", "CXCR4+ NK", "CD68- Monocytes", "CD8A T", "CD14- INFG- Monocytes", "B cells", "CXCR4- NK", "Non classical monocytes", "Monocytes", "RUNX3+ NK", "Myeloid DC", "Low-density basophils", "Effector Memory T", "Th1")
response <- c("NR", "R")

Sco@meta.data <- data.frame(label = Sco$response, cell_type = as.integer(Sco$seurat_clusters))
Sco@meta.data[,1] <- sapply(Sco@meta.data[,1], function(x) response[x+1])
Sco@meta.data[,2] <- sapply(Sco@meta.data[,2], function(x) cell_types[x])

head(Sco@meta.data)
```

Now we can run the Augur algorithm with default parameters

```{r, eval = FALSE}
augur = Augur::calculate_auc(Sco, n_threads = 8)

save(augur, file = "/Users/phillipnicol/Desktop/local_files/data/scDiffPopDATA/Augur/augur_res.RData")
```

The above chunk takes around 20 minutes to complete on my laptop.


## Results

Let's take a look at what we got. First, we load the list from our files.

```{r}
load("/Users/phillipnicol/Desktop/local_files/data/scDiffPopDATA/Augur/augur_res.RData")
```

Now we take a look at the AUC 

```{r}
augur$AUC
```

Note that the algorithm cut out low-density basophils, since there are only 14 responding cells which have this cell type and the default settings of their algorithm requires at least $20$ cells in each class (maybe this is interesting). Let's also plot it 


```{r}
library(ggplot2)

#prepare colors
library(Polychrome)
alfa <- alphabet.colors(14)

#plot
df <- as.data.frame(augur$AUC)
p <- ggplot(data=df, aes(x=cell_type, y = auc)) + geom_bar(stat = "identity", fill = alfa)
p <- p + xlab("Cell type") + ylab("AUC (Augur)") + coord_cartesian(ylim=c(0.5,0.75))
p <- p + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
plot(p)

#save the plot
png(file = "/Users/phillipnicol/Desktop/local_files/data/scDiffPopDATA/Augur/auc_pbmc.png")
```


